---
title: "FARS Vignette"
author: "George Harrelson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FARS Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Filename Creation

The *make_filename()* function creates the filename of a dataset based on the year you input. The output from this
function can be used by the *fars_read()* function.

Note: There is no error checking here, so it is possible to put in a year that will not end up reading in a dataset.

### Parameters

|**Parameters** |**Description**                                                 |   
|---------------|----------------------------------------------------------------|
|year| A numeric input used to define a filename you want to read in.|

### Example

```{r eval = FALSE}
make_filename(2015)
```

## Read Data In

The *fars_read()* function reads in data from the Fatality Analysis Reporting System (FARS). The input is a filename.
If the filename cannot be found then an error will be returned, otherwise the file will be read in.

### Parameters

|**Parameters** |**Description**                                   |
|---------------|--------------------------------------------------|
|filename| A string input containing the filename you want read in.|

### Example

```{r eval = FALSE}
fars_read("accident_2015.csv.bz2")
```

## Summarize Data by Month/Year

The *fars_read_years()* function reads in multiple years of data and summarizes accidents by month and year. This function will return a list for each year of data read in. Invalid year inputs will generate a warning message, but as long as there is at least one valid year then this function will return a list.

### Parameters

|**Parameters** |**Description**                                                  |
|---------------|-----------------------------------------------------------------|
|years| A list of numeric values for the years you want to summarize FARS data for|

### Example

```{r eval = FALSE}
fars_read_years(c(2014,2015))
```

## Summarize Data by Year

The *fars_summarize_years()* function will summarize the specified data by month and year in a table. Invalid year inputs will generate a warning message, but as long as there is at least one valid year input then this function will return a table.

### Parameters

|**Parameters** |**Description**                                                  |
|---------------|-----------------------------------------------------------------|
|years| A list of numeric values for the years you want to summarize FARS data for|

### Example

```{r eval = FALSE}
fars_summarize_years(c(2014,2015))
```

## State Maps

The *fars_map_state()* function will output a graphical map showing where within a state fatal accidents occur. State numbers outside of the valid range (1-56) will generate an error. States with no accidents to plot will generate a warning message. This function will also generate errors for invalid year inputs.

### Parameters

|**Parameters** |**Description**                                                  |
|---------------|-----------------------------------------------------------------|
|state.num| A numeric value that represents the state you want to plot data for   |
|years| A list of numeric values for the years you want to summarize FARS data for |

### Example

```{r eval = FALSE}
fars_map_state(45, 2015)
```
